Eliminating prior-bias from sparse-projection tomographic reconstructions

### 2020-06-02 paper/tci.tex
added SSIM table for Fig.8 (2D reconstruction of potato)

In addition, when some extra information about the object's structure is known~\cite{PICCS,cardiacPICCS,lubner2011,pirple,mota2017}, it is beneficial in further reducing the number of measurements.  These prior-based techniques use information from previously scanned data (`object-prior') of the same object or a similar one, and utilize it to reconstruct new volumes from very few sets of additional measurements. However, the above methods have the following two limitations: (1) the prior information may potentially overwhelm the essential details that appear in the test, (2) there is the key issue of choice of \textit{a} particular previously scanned object among the many previously acquired scans. A large part of this paper is dedicated to alleviating the former limitation, which is an important issue that has so far been overlooked in the literature on tomography. The latter limitation was relaxed in~\cite{liu2016,Xu2012} by building dictionary-based priors from multiple previously scanned objects. However, as reported in~\cite{my_dicta_paper}, dictionary priors are not as accurate and fast as global eigenspace priors.  Global eigenspace priors are better able to exploit the similarity of a test volume to object-prior, by assuming that the new test volume lies within the space spanned by the eigenvectors of the multiple representative previously scanned objects. Fig.~\ref{fig:diff_methods} shows the advantage of combining the global prior with CS.
\subsection{\textbf{Contributions}}
\label{sec:contributions}
In this work (Fig.~\ref{fig:prior_overview}), we focus on further reducing the number of measurements, with  particular emphasis on longitudinal studies. For example, we consider medical datasets which consist of multiple CT scans being taken during a radio-frequency ablation procedure~\cite{Dong2015}. The process consists of inserting a needle probe into a patient's body. When the needle reaches the tumor site, an electric current is passed to burn the tumor. Repeated CT scans of the patient need to be acquired in order to visualize the movement of the needle and to ensure that it is reaching the appropriate target. A few initial densely-sampled scans are used as object-prior to help the physician know the position of the needle. The later scans are used to reveal the exact changes during and after burning of the tumor (ablation).  In this context, we demonstrate the combined use of CS with the global prior, in two flavors: the vanilla (unselective) and selective global prior-based reconstruction. The choice of the number of measurement views and the type of reconstruction - unselective or selective, is driven by the goal of the procedure.

@@

%-----------------------------potato 2D weights
\begin{figure}[!h]
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_fbp30.png}
        \caption{FBP}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_cs_dct30.png}
        \caption{CS-DCT}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_cs_wavelet30.png}
        \caption{CS-Haar}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_art30.png}
        \caption{ART}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_sart30.png}
        \caption{SART}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_sirt30.png}
        \caption{SIRT}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_all_methods30.png}
        \caption{combined}
     \end{subfigure}
      \caption{Weights-maps (corresponding to the difference between pilot reconstruction of the image in the last sub-figure of Fig.~\ref{fig:object-prior_test_potato_A1} and its projection onto the eigenspace $V_{low}$) constructed using different reconstruction methods, as specified in Eq.~\ref{eq:weightsEq}. The weights-maps are different because the reconstruction artefacts of the new structures in test image will be different for every reconstruction method used, as seen in Fig.~\ref{fig:reconstructions_diff_methods}. }
\label{fig:weights_map_2Dpotato}
\end{figure}
%-----------------------------potato 2D weights
\begin{figure}[!h]
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/fbp.png}
        \caption{FBP}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/cs_dct.png}
        \caption{CS-DCT}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/cs_wavelet.png}
        \caption{CS-Haar}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/art.png}
        \caption{ART}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/sart.png}
        \caption{SART}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/sirt.png}
        \caption{SIRT}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weighted_pca_all_methods30.png}
        \caption{combined}
     \end{subfigure}
      \caption{(a)-(f):Different reconstructions of~\ref{fig:potato_test}. The magnitude and sharpness of the artefacts is different for each method. (g) Selective-prior method that combines weights-map information from all other methods.} 
\label{fig:reconstructions_diff_methods}
\end{figure}

The changes and new structures present in the test data will generate different artifacts for different reconstruction techniques. These artifacts would not be captured by reconstructions of the object-prior since the underlying new changes and structures may be absent in all of the previously scanned objects. We aim to let the weights be independent of the type of artifact. Hence, we use a combination of different reconstruction techniques to generate different types of eigenspaces and combine information from all of them to compute weights. To illustrate the benefit of this method, we first performed 2D reconstruction of a test slice from the potato dataset (please refer Sec.~\ref{Sec:potato} for details of the dataset) Fig.~\ref{fig:object-prior_test_potato_A1} shows the test and template slices. Fig.~\ref{fig:weights_map_2Dpotato}  shows the weights-maps generated using Eq.~\ref{eq:weightsEq} by various reconstruction methods. It can be seen that the weights are low in the region of the new change in test data. Because all the iterative methods are computationally expensive, we chose only FBP and CS-DCT for computing weights-maps for all 3D reconstructions.

>>>>

In addition, when some extra information about the object's structure is known~\cite{PICCS,cardiacPICCS,lubner2011,pirple,mota2017}, it is beneficial in further reducing the number of measurements.  These prior-based techniques use information from previously scanned data (`object-prior') of the same object or a similar one, and utilize it to reconstruct new volumes from very few sets of additional measurements. However, the above methods have the following two limitations: (1) the prior information may potentially overwhelm the essential details that appear in the test, (2) there is the key issue of choice of \textit{a} particular previously scanned object among the many previously acquired scans. A large part of this paper is dedicated to alleviating the former limitation, which is an important issue that has so far been overlooked in the literature on tomography. The latter limitation was relaxed in~\cite{liu2016,Xu2012} by building dictionary-based priors from multiple previously scanned objects. However, as reported in~\cite{my_dicta_paper}, dictionary priors are not as accurate and fast as global eigenspace priors.  Global eigenspace priors are better able to exploit the similarity of a test volume to object-prior, by assuming that the new test volume lies within the space spanned by the eigenvectors of the multiple representative previously scanned objects. Fig.~\ref{fig:diff_methods} shows the advantage of combining the global prior with CS.
\subsection{\textbf{Contributions}}
\label{sec:contributions}
In this work (Fig.~\ref{fig:prior_overview}), we focus on further reducing the number of measurements, with  particular emphasis on longitudinal studies. For example, we consider medical datasets which consist of multiple CT scans being taken during a radio-frequency ablation procedure~\cite{Dong2015}. The process consists of inserting a needle probe into a patient's body. When the needle reaches the tumor site, an electric current is passed to burn the tumor. Repeated CT scans of the patient need to be acquired in order to visualize the movement of the needle and to ensure that it is reaching the appropriate target. A few initial densely-sampled scans are used as object-prior to help the physician know the position of the needle. The later scans are used to reveal the exact changes during and after burning of the tumor (ablation).  In this context, we demonstrate the combined use of CS with the global prior, in two flavors: the vanilla (unselective) and selective global prior-based reconstruction. The choice of the number of measurement views and the type of reconstruction -- unselective or selective, is driven by the goal of the procedure.

@@

%-----------------------------potato 2D weights
\begin{figure}[!h]
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_fbp30.png}
        \caption{FBP}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_cs_dct30.png}
        \caption{CS-DCT}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_cs_wavelet30.png}
        \caption{CS-Haar}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_art30.png}
        \caption{ART}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_sart30.png}
        \caption{SART}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_sirt30.png}
        \caption{SIRT}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_all_methods30.png}
        \caption{combined}
     \end{subfigure}
      \caption{Weights-maps (corresponding to the difference between pilot reconstruction of the image in the last sub-figure of Fig.~\ref{fig:object-prior_test_potato_A1} and its projection onto the eigenspace $V_{low}$) constructed using different reconstruction methods, as specified in Eq.~\ref{eq:weightsEq}. The weights-maps are different because the reconstruction artefacts of the new structures in test image will be different for every reconstruction method used, as seen in Fig.~\ref{fig:reconstructions_diff_methods}. }
\label{fig:weights_map_2Dpotato}
\end{figure}
%-----------------------------potato 2D weights
\begin{figure}[!h]
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/fbp.png}
        \caption{FBP}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/cs_dct.png}
        \caption{CS-DCT}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/cs_wavelet.png}
        \caption{CS-Haar}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/art.png}
        \caption{ART}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/sart.png}
        \caption{SART}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/sirt.png}
        \caption{SIRT}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weighted_pca_all_methods30.png}
        \caption{combined}
     \end{subfigure}
      \caption{(a)-(f):Different reconstructions of~\ref{fig:potato_test}. The magnitude and sharpness of the artefacts is different for each method. (g) Selective-prior method that combines weights-map information from all other methods. The SSIM of these reconstructions is shown in Table.~\ref{table:potato_2D_ssim}.} 
\label{fig:reconstructions_diff_methods}
\end{figure}

The changes and new structures present in the test data will generate different artifacts for different reconstruction techniques. These artifacts would not be captured by reconstructions of the object-prior since the underlying new changes and structures may be absent in all of the previously scanned objects. We aim to let the weights be independent of the type of artifact. Hence, we use a combination of different reconstruction techniques to generate different types of eigenspaces and combine information from all of them to compute weights. To illustrate the benefit of this method, we first performed 2D reconstruction of a test slice from the potato dataset (please refer Sec.~\ref{Sec:potato} for details of the dataset) Fig.~\ref{fig:object-prior_test_potato_A1} shows the test and template slices. Fig.~\ref{fig:weights_map_2Dpotato}  shows the weights-maps generated using Eq.~\ref{eq:weightsEq} by various reconstruction methods. It can be seen that the weights are low in the region of the new change in test data. Because all the iterative methods are computationally expensive, we chose only FBP and CS-DCT for computing weights-maps for all 3D reconstructions.  
%---------------------------------------
\begin{table}[!h]
\caption{SSIM of the reconstructions shown in Fig.~\ref{fig:reconstructions_diff_methods}:(a)-(g). The SSIM of the ground-truth (ideal reconstruction) is 1, and these are computed within the red RoI of the test image shown in Fig.~\ref{fig:potato_test}.}
\begin{tabular}{|l|c|c|c|c|c|c|c|c|}
\hline
 Fig.~\ref{fig:reconstructions_diff_methods}  & \textbf{(a)} & \textbf{(b)} & \textbf{(c)} & \textbf{(d)} & \textbf{(e)} & \textbf{(f)} &  \textbf{(g)} \\\hline
\textbf{red RoI}  & 0.502 & 0.597  & 0.453 & 0.474 & 0.565 & 0.523 & \textcolor{red}{0.759} \\ \hline
\end{tabular}
\label{table:potato_2D_ssim}
\end{table}
%-------------------------------------------------
